---
title: "Fast Data Processing in R"
author: "Nicholas Potter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Sometimes we run into **Medium Data**, data that are too big to be easily used on our computers but still small enough that they *can* be used, with some special tools.

###Data IO
Reading in raw csv files of larger than 500MB can take some time. And the entire dataset must be stored in memory. There are a few solutions to speed this up, with various trade offs.

####Loadable in Sections
If the data you need loaded at a given point is small enough to fit in RAM, then `data.table` is a great package. `data.table` includes `fread`, which provides much faster read times, and you can select specific columns rather than the entire data set. 

``` {r eval = FALSE}
library(data.table)
#select 100000 rows, just columns 1, 4, and 24
fread("<data file>", select = c(1,4,4), nrows = 100000)
```

Combined with **sampling**, you can probably do much of the exploratory analysis without ever loading the full data set. 

####Selectable
If the data are too big to hold and analyze in RAM, or you want more specific selection criteria than available when reading a file using `fread`, then either flat files or databases work well.

[SQLite](https://sqlite.org/) is a simple free database that is easy to use. Importing a csv file into sqlite is as simple as writing an sqlite script.

```
.mode csv
.headers on
.import "<csv file>" table_name
```

Then with `dplyr` you can select data from the sqlite database easily:

``` {r eval = FALSE}
db <- dplyr::src_sqlite("path/to/sqlite_database")
df < tbl(db, "table_name") %>%
  filter(tbl, year>2000, income>40000) %>%
  collect()
```

This returns just the selected data, while allowing the selection of a specific subset of data. There are numerous SQL database options, and much that you can do with `dplyr`'s interface for selecting, sorting, and summarizing data.

Another option is `ff` and `bit`, which keeps the data in files on the hard drive and allows for much of the same indexing and selecting behavior as `data.frames`. It makes extensive use of chunking. A csv file can be read into ffdf files and saved to an ffdf object by:

``` {r eval = FALSE}
library(ff)
test <- read.csv.ffdf(file = "<csv_file>",
                      first.rows = 10000,
                      next.rows = 50000)

ffsave(test,
       file = "<file_name>",
       compress = TRUE,
       rootpath = ".")
```

The data can be indexed and sorted without loading into memory, or read into memory just in sections that you want:

``` {r eval = FALSE}
ffload("data/test", overwrite = TRUE, rootpath = "data/")

#Randomly sort the data
N <- dim(test)
s <- runif(N)
test1 <- ffdf(test[order(r), c("col1", "col2", "col3")])

#Read ffdf formatted data into memory:
dat <- test[1:100000, ] #is a data.frame

#Processing can happen in chunks:
library(bit)
for (i in chunk(1, N, 10000)) {
  #Do some processing to the data
}
```

###Sampling
Sampling can be hugely helpful, 
You can benchmark the selection yourself
In benchmarks of 


Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
