---
title: "A First Analysis with XGboost"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A First Analysis with XGboost}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[`xgboost`](http://www.cran.r-project.org/web/packages/xgboost/index.html) is a fast implementation of gradient boosting. As a first approach we implement a random forest using just the records that have a radar reflectivity measure (`Ref`). The model here is just a linear model of reflectivity, which of course doesn't work very well, but gives an idea of how to set up xgboost.

``` {r eval = FALSE}
library(ff)
library(bit)
library(xgboost)
library(kgRainPredictR)
ffload("data/train", overwrite = TRUE, rootpath = "data/")

#Filter NA values for Ref
N <- dim(train)[1]
b <- bit(N)
for (i in chunk(1,N,10000)) {
  b[i] = is.na(train$Ref[i])
}
df_tr <- as.data.frame(train[!b, c("Id", "minutes_past", "Ref", "Expected")]) %>% 
  filter(Expected <= 69) %>%
  group_by(Id) %>%
  summarize(ref_mean = mean(time_difference(minutes_past) * Ref, na.rm = TRUE), 
            Expected = max(Expected))


labels <- df_tr$Expected

bst <- xgboost(data = as.matrix(df_tr[,c("Id", "ref_mean")]),
               label = labels, 
               max.depth = 3, 
               eta = .7, 
               nround = 10,
               objective = "reg:linear",
               missing = NA,
               base_score = 0,
               verbose = 1)

xgb.save(bst, 'bst_no_na.save')

#clean up
rm(df_tr)
rm(b)
rm(labels)
```

At this point we have a trained model and have saved it to a file. We've also removed our loaded data because we need to make everything available that we can for prediction.

``` {r eval = FALSE}
#Now load the test data
ffload("data/test", overwrite = TRUE, rootpath = "data/")
df_te <- as.data.frame(test[,c("Id", "minutes_past", "Ref")]) %>%
  group_by(Id) %>%
  summarize(ref_mean = mean(time_difference(minutes_past) * Ref, na.rm = TRUE))

res <- data.frame(
  Id = df_te$Id,
  Expected = predict(bst, as.matrix(df_te), missing = NA)
)

res <- group_by(res, Id) %>% summarize(Expected = mean(Expected))
createSubmission(res, "submission.csv")


```
